---
title: "Bio3 JB"
author: "Jason Baer"
date: "10/18/2019"
output: html_document
---

# Question 1 - read.pdb() function returns a list
# Question 2 - trim.pdb() function trims the function by removing remarks and many xyz atoms
# Question 3 - the input parameters top and bot would turn off the black and grey rectangles , which tells per-residue numeric vectors for a given protein structure
# Question 4 - a clustered dendrogram to show similarity in B-factor trends would be more useful because you can see the similarity on the same plot. 
# Question 5 - s1.b and s3.b are more similar in their B-factor trends which can be quantified in the cluster dendrogram shown below. rbind first compares based on B-factors (limits analysis to that row), then dist determines the distance between each, and then hclust creates a clustered hierarchy with the three variables. 
# Question 6 - shown below 

```{r}
library(bio3d)
pdb_reader <- function(x) {
  s <- read.pdb(x) # reads pdb file of protein
s.chainA <- trim.pdb(s, chain="A", elety="CA") # trims pdb list to only look at one side chain
s.b <- s.chainA$atom$b # identifies values of b factor on specific side chain

plotb3(s.b, sse=s.chainA, typ="l", ylab="Bfactor") # produces a plot to show b factor values on each residue

}
pdb_reader("4AKE")
```

```{r}
library(bio3d)
pdb_reader <- function(x) {
  s <- read.pdb(x) # reads pdb file of protein
s.chainA <- trim.pdb(s, chain="A", elety="CA") # trims pdb list to only look at one side chain
s.b <- s.chainA$atom$b # identifies values of b factor on specific side chain

plotb3(s.b, sse=s.chainA, typ="l", ylab="Bfactor") # produces a plot to show b factor values on each residue

}
pdb_reader("1AKE")
```
```{r}
library(bio3d)
pdb_reader <- function(x) {
  s <- read.pdb(x) # reads pdb file of protein
s.chainA <- trim.pdb(s, chain="A", elety="CA") # trims pdb list to only look at one side chain
s.b <- s.chainA$atom$b # identifies values of b factor on specific side chain

plotb3(s.b, sse=s.chainA, typ="l", ylab="Bfactor") # produces a plot to show b factor values on each residue

}
pdb_reader("1E4Y")
```

```




