---
title: "Class 14"
author: "Jason Baer"
date: "11/13/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r eval = FALSE}
intall.packages("BiocManager")
BiocManager::install("DESeq2")
library(BiocManager)
library(DESeq2)
```


```{r}
counts <- read.csv("airway_scaledcounts.csv", stringsAsFactors = FALSE)
counts

```

```{r}
metadata <- read.csv("airway_metadata.csv", stringsAsFactors = FALSE)
metadata

```

Q1: How many genes are in the dataset?

```{r}
nrow(counts)
```

How many experiments?

```{r}
ncol(counts) - 1
```

Lets make sure metadata id col matches the colnames of counts

```{r}
all(colnames(counts)[-1] == metadata$id )
```


Q2: How many control cell lines do we have?

```{r}
metadata$dex == "control"
```

First we want to access the columns of countData that are control and then find their mean.

```{r}
control.id <- metadata[metadata$dex == "control", ]$id
control.id
```


```{r}
control.mean <- rowSums(counts[, control.id])/length(control.id)
names(control.mean) <- counts$ensgene
head(control.mean)
```


Then we can do the same for treated samples

```{r}
treated.id <- metadata[metadata$dex == "treated", ]$id
treated.id
```


```{r}
treated.mean <- rowSums(counts[ ,treated.id]) / length(treated.id)
names(treated.mean) <- counts$ensgene
head(treated.mean)
```

Combine meancount data into a dataframe

```{r}
meancounts <- data.frame(control.mean, treated.mean)
meancounts
```

Plot control vs treated

```{r}
plot(meancounts$control.mean, meancounts$treated.mean, xlab = "Control Mean", ylab = "Treated Mean")
```

Plot both of these on a log scale

```{r}
plot(meancounts$control.mean, meancounts$treated.mean, xlab = "Control Mean", ylab = "Treated Mean", log = "xy")
```

Fold change. negative numbers mean fold change is going down, positive mean it is going up. -inf means negative infinity (taking log of zero), so genes that have zero expression. 

```{r}
meancounts$log2fc <- log2(meancounts[,"treated.mean"]/meancounts[,"control.mean"])
head(meancounts)
```

Lets filter out genes with no expression and NaN

```{r}
zero.vals <- which(meancounts[,1:2]==0, arr.ind=TRUE)

to.rm <- unique(zero.vals[,1])
mycounts <- meancounts[-to.rm,]
head(mycounts)
```


Test for finding zero entries

```{r}
x <- data.frame(happy = c(5,6,0,0), sad = c(0,5,5,0))
x == 0
which(x == 0)

which(x == 0, arr.ind = TRUE)
```

A common threshold used for calling something differentially expressed is a log2(FoldChange) of greater than 2 or less than -2. Let’s filter the dataset both ways to see how many genes are up or down-regulated.

```{r}
up.ind <- mycounts$log2fc > 2
down.ind <- mycounts$log2fc < (-2)
length(which(up.ind, arr.ind = TRUE))
length(which(down.ind, arr.ind = TRUE))
```

